---
title: "Automated incidence rate of diagnosis data"
author: "Jinhee Chang"
date: "9/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal of this code: Automate the incidence rate of diagnosis data

#master5 data includes patient ID, Disease code, whether this is ICD9/ICD10 based, the date patient got diagnosed, whether the diagnosis was main or secondary, patient's birthday, Age patient got diagnosed, meaning of the diagnosis, etc - Total: 264,069 patients, 3,003,823 diagnosis data

#All data was retrieved from UK biobank (https://www.ukbiobank.ac.uk/) and was downloaded using Stampede2. I created a custom file (master5) by merging multiple datasets for my convenience in the past.

```{r}
# Load the data that I am going to use
master6 <- readRDS("master_6.rds")

# Check how this data looks like
head(master6)

# Load ICD10_Code data
ICD10_Code <- readr::read_tsv('coding19.tsv')
ICD9_Code <- readr::read_tsv('coding87.tsv')
```

```{r}
## Don't forget to download this function
library(tidyr)
library(dplyr)
library(data.table)

sum(is.na(master6$Diagnosed_age)) #1230776 NA exist
# Delete all the rows that have NA on Diagnosed age column
master6 <- master6[complete.cases(master6), ] #3003823 rows left

# How many patient records are unique
length(unique(master6$eid)) #264069 patients total
```
## Check the age distribution of the total unique patients in master5 file

```{r}
#hist(unique_master5$Diagnosed_age, breaks = 50)
hist(master6$Diagnosed_age, breaks = 50)

# Check the minimum age and maximum age of the unique_master5
min(master6$Diagnosed_age) # 27.93699
max(master6$Diagnosed_age) # 80.87397
```

##Make incidence rate using Total Disease age distribution
```{r}
# Total Disease incidence rate: totalage_freq
totalage_freq <- as.data.frame(table(floor(master6$Diagnosed_age)))
colnames(totalage_freq) <- c("Age at Diagnosis", "Frequency")
totalage_freq$`Incidence Rate` <-
  totalage_freq$Frequency/sum(totalage_freq$Frequency)
totalage_freq$category <- rep(paste0("Total Disease", "(n=",
                                     sum(totalage_freq$Frequency), ")"),
                              nrow(totalage_freq))
totalage_freq$`Age at Diagnosis` <- as.numeric(as.character(totalage_freq$`Age at Diagnosis`))

# Check how this looks like
totalage_freq

#saveRDS(totalage_freq, file = "totalage_freq.rds")
```
## G30 Alzheimer's disease (AD) - No AD in ICD9 ("3310")
G30.0 Alzheimer's disease with early onset    95
G30.1 Alzheimer's disease with late onset     41
G30.8 Other Alzheimer's disease               90
G30.9 Alzheimer's disease, unspecified        815

## I50 Heart failure (CHF) - 17 more in ICD9 ("428")

#master5 %>% filter(str_detect(Disease_Code, "^428"))

I50.0 Congestive heart failure          3607
I50.1 Left ventricular failure          5398
I50.9 Heart failure, unspecified        2532

## J44 Other chronic obstructive pulmonary disease (COPD) - 7 more in ICD9 ("491")

#master5 %>% filter(str_detect(Disease_Code, "^491"))

J44.0 Chronic obstructive pulmonary disease with acute lower respiratory infection       3703
J44.1 Chronic obstructive pulmonary disease with acute exacerbation, unspecified         1707
J44.8 Other specified chronic obstructive pulmonary disease                              1042
J44.9 Chronic obstructive pulmonary disease, unspecified                                11734

## I21 Acute myocardial infarction (Acute MI) - 168 more in ICD9 ("410")

#master5 %>% filter(str_detect(Disease_Code, "^410"))

I21.0 Acute transmural myocardial infarction of anterior wall       2283
I21.1 Acute transmural myocardial infarction of inferior wall       2785
I21.2 Acute transmural myocardial infarction of other sites         341
I21.3 Acute transmural myocardial infarction of unspecified site    171
I21.4 Acute subendocardial myocardial infarction                    2168
I21.9 Acute myocardial infarction, unspecified                      4789

## I22 Subsequent myocardial infarction (Subsequent MI)
I22.0 Subsequent myocardial infarction of anterior wall             117
I22.1 Subsequent myocardial infarction of inferior wall             205
I22.8 Subsequent myocardial infarction of other sites               90
I22.9 Subsequent myocardial infarction of unspecified site          545

## I64 Stroke, not specified as haemorrhage or infarction  1432 - 11 more in ICD9 ("434")

#master5 %>% filter(str_detect(Disease_Code, "^434"))

## E11 Non-insulin-dependent diabetes mellitus (Diabetes Type II) - 104 more in ICD9 ("2500")

#master5 %>% filter(str_detect(Disease_Code, "^2500"))

E11.0 With coma                                       105
E11.1 With ketoacidosis                               220
E11.2 With renal complications                        394
E11.3 With ophthalmic complications                   1861
E11.4 With neurological complications                 815
E11.5 With peripheral circulatory complications       506
E11.6 With other specified complications              153
E11.7 With multiple complications                     10
E11.8 With unspecified complications                  86
E11.9 Without complications                           28852

## E14 Unspecified diabetes mellitus (Unspecified Diabetes) - 5 more in ICD9 ("2509")

#master5 %>% filter(str_detect(Disease_Code, "^2509"))

E14.0 With coma                                       14
E14.1 With ketoacidosis                               34
E14.2 Withrenal complications                         31
E14.3 With ophthalmic complications                   389
E14.4 With neurological complications                 49
E14.5 With peripheral circulatory complications       27
E14.6 With other specified complications              25
E14.7 With multiple complications                     3
E14.8 With unspecified complications                  35
E14.9 Without complications                           3421

## Total base disease code
```{r}
source("Automated_functioncode.R")

base_diseasecode <- c("G30", "E11", "E14", "I21", "I22", "I50", "I64", "J44")

for (i in (1:length(base_diseasecode))) {
  dataplotting(original_ICD9(base_diseasecode[i]), 2)
  dataplotting(original(base_diseasecode[i]), 2)
  dataplotting(normalized(base_diseasecode[i]), NULL)
}
```

## From Geneatlas

```{r}
geneatlas <- read.table("geneatlas.txt", header = TRUE)
nrow(geneatlas) # 157
length(which(duplicated(geneatlas$ICD10))) # 46
geneatlas <- geneatlas[!duplicated(geneatlas$ICD10), ]

duplicated_code <- c("G30", "E11", "E14", "I21", "I22", "I50", "I64", "J44")

sum(geneatlas$ICD10 %in% duplicated_code) #7

geneatlas_ICD10 <- geneatlas$ICD10[! geneatlas$ICD10 %in% duplicated_code]
geneatlas_ICD10 <- geneatlas_ICD10[!is.na(geneatlas_ICD10)] # total 103
```

```{r}
for (i in (1:length(geneatlas_ICD10))) {
  dataplotting(original_ICD9(geneatlas_ICD10[i]), 2)
  dataplotting(original(geneatlas_ICD10[i]), 2)
  dataplotting(normalized(geneatlas_ICD10[i]), NULL)
}
```

