---
title: "Automated_deathfile"
author: "Jinhee Chang"
date: "10/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ultimate Goal: Create death incidence rate and frequency
library(dplyr)
library(readr)

```{r}
# Load the data that I am going to use
deathfile <- readRDS("Deathfile.rds")

ICD10_Code <- readr::read_tsv('coding19.tsv')
#unique_diseasecode <- master5[which(!duplicated(master5$Disease_Code)), ]
#unique_diseasecode %>% mutate(across(where(is.factor), as.character)) -> unique_diseasecode

for (i in 1:nrow(deathfile)) {
  if (deathfile$Diseasecode[i] %in% ICD10_Code$coding) {
    deathfile$Descriptioncauseofdeath[i] <- ICD10_Code$meaning[ICD10_Code$coding == deathfile$Diseasecode[i]]
  } else {
    deathfile$Descriptioncauseofdeath[i] <- "Unspecified Death"
  }
}

saveRDS(deathfile, file = "Deathfile.RDS")

# Check how this data looks like
head(deathfile)
```

## Calculate death incidence rate as a whole
library(tidyr)
library(dplyr)
library(data.table)

```{r}
# Check if there is any NA in ageatdeath
which(is.na(deathfile$Ageatdeath)) # none

# How many death record are unique
length(unique(deathfile$eid)) #20465
unique_death <- deathfile[which(!duplicated(deathfile$eid)), ]
checkdeath <- table(unique(unique_death$eid) %in% unique(unique_primarydeath$eid)) # 33 NA Diseasecode
unique_death

# Unique primary death record
unique_primarydeath <- subset(deathfile, Causeofdeath == "Primary")
unique_primarydeath # 20473 rows

# Duplicate primary death
length(which(duplicated(unique_primarydeath$eid))) # 41
length(unique(unique_primarydeath$eid)) #20432
```

## Check the age range of the unique patients in "deathrate" (using unique_death)

```{r}
# Check the age distribution of unique_death patients
hist(unique_death$Ageatdeath, breaks = 50)

# Check the minimum age and maximum age of the data
min(unique_death$Ageatdeath) # 40
max(unique_death$Ageatdeath) # 81
```

## Draw line graph of unqiue death patients
library(ggplot2)

```{r}
deathage_freq <- as.data.frame(table(unique_death$Ageatdeath))
colnames(deathage_freq) <- c("Age at Death", "Frequency")
deathage_freq$`Death Rate` <- deathage_freq$Frequency/sum(deathage_freq$Frequency)
deathage_freq$`Age at Death` <- as.numeric(as.character(deathage_freq$`Age at Death`))
# which frequency of deathage is over 500
deathage_freq[which(deathage_freq$Frequency > 500),] # min 61, max 76

ggplot(data = deathage_freq, aes(x = `Age at Death`,
                                  y = log10(`Death Rate`)), las = 3)+
  geom_line()+
  scale_x_continuous(limits = c(61, 76), n.breaks = 10)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(title= "Death Frequency (only over 500)", y="log10(Death Rate)", x = "Age Range")+
  theme_bw()
```

```{r}
test_death <- function(code_num, freq = 100, age_min = 40, age_max = 70) {
  
  result <- deathfile[grep(code_num, deathfile$Diseasecode), ]
  subset_name <- result %>% distinct(Diseasecode)
  
  subset_list <- list()
  for (i in 1:nrow(subset_name)) {
    result_subset <- deathfile[grep(subset_name[i,], deathfile$Diseasecode), ]
    age_freq <- as.data.frame(table(result_subset$Ageatdeath))
    colnames(age_freq) <- c("Age at Death", "Frequency")
    
    if (sum(age_freq$"Frequency") > freq) {
      age_freq$`Death Rate` <- age_freq$Frequency/sum(age_freq$Frequency)
      age_freq$category <- rep(paste0(result_subset$Descriptioncauseofdeath[1],
                                    "(n=", nrow(result_subset), ")"),
                               nrow(age_freq))
      subset_list[[i]] <- age_freq
      }
  }
  
  totaldataframe <- rbindlist(subset_list)
  
  age_freq <- as.data.frame(table(result$Ageatdeath))
  colnames(age_freq) <- c("Age at Death", "Frequency")
  age_freq$`Death Rate` <- age_freq$Frequency/sum(age_freq$Frequency)
  age_freq$category <- rep(paste0("Total Death", "(n=", nrow(result), ")"),
                           nrow(age_freq))

  totaldataframe <- rbind(totaldataframe, age_freq)
  totaldataframe$`Age at Death` <- as.numeric(as.character(totaldataframe$`Age at Death`))
  totaldataframe$category <- as.factor(totaldataframe$category)
  
  # Draw line plot
  plotdata<- ggplot(data = totaldataframe, aes(x = `Age at Death`,
                                  y = log10(`Death Rate`), color = category,
                                  group = category), las = 3)+
  geom_line()+
  scale_x_continuous(limits = c(age_min, age_max), n.breaks = 10)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(y = "log10(Death Rate)", x = "Age at Death")+
  theme_bw()+
  theme(legend.position = "right", 
        legend.direction = "vertical",
        legend.background = element_rect(color = "steelblue", linetype = "solid"),
        legend.text = element_text(size = 8))
  #+guides(col = guide_legend(ncol = 2))
    
  print(plotdata)
  
  png(paste0(code_num, "_deathplot.png"), res = 80, width = 1000, height = 600)
  print(plotdata)
  dev.off() 
  
  return(age_freq)

}
```

## G30 Alzheimer's disease

```{r}
# AD
test_death("G30")
```

## I50 Heart failure-

```{r}
# CHF
test_death("I50")
```

## J44 Other chronic obstructive pulmonary disease-

```{r}
# COPD
test_death("J44")
```

## I21 Acute myocardial infarction-

```{r}
# Acute MI
test_death("I21")
```

## I22 Subsequent myocardial infarction - NONE
## I64 Stroke, not specified as haemorrhage or infarction 

```{r}
# Stroke
test_death("I64")
```

## E11 Non-insulin-dependent diabetes mellitus-

```{r}
# Diabetes Type II
test_death("E11")
```

## E14 Unspecified diabetes mellitus-

```{r}
# Unspecified Diabetes 
test_death("E14")
```
