---
title: "Phecode_UKB"
author: "Jinhee Chang"
date: "12/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Read the file that is going to be the base data
```{r}
phecode_ukb <- readr::read_tsv('phecode-ukbb-agg.tsv')
phecode_ukb$phecode <- sprintf("%.2f", phecode_ukb$phecode)
```
##Dataplotting for normalized_phecode
```{r}
source("Automated_functioncode.R")

for (i in (1:length(phecode_ukb$phecode))) {
  dataplotting(normalized_phecode(phecode_ukb$phecode[i]), NULL)
}
```

##preprocess_clustering is the result of function "clustering_preprocess" and this part is omitted in the chunk
##Make preprocess_clustering_60cut 
```{r}
source("Automated_functioncode.R")

preprocess_clustering_60cut <- subsetting_cluster_result(preprocess_clustering, 60)
```

##Drawing ComplexHeatmap using preprocess_clustering and preprocess_clustering_60cut
```{r}
library("ComplexHeatmap")

library(circlize)
col_fun = colorRamp2(c(0, 1, 2, 4), c("white", "green", "yellow", "red"))

Heatmap(preprocess_clustering, cluster_rows = FALSE, col = col_fun,
        column_split = 8)
Heatmap(preprocess_clustering_60cut, cluster_rows = FALSE, col = col_fun,
        column_split = 8)
```
##SOM drawing (2 ways)
```{r}
library(kohonen)

allres <- som(t(preprocess_clustering), grid = somgrid(5, 5))
plot(allres)

# reverse color ramp
colors <- function(n, alpha = 1) {
    rev(heat.colors(n, alpha))
}

plot(allres, type = "counts", palette.name = colors, heatkey = TRUE)

som_grid <- somgrid(xdim = 5, ydim=5, topo="hexagonal")

som_model <- som(t(preprocess_clustering), 
    grid=som_grid, 
    rlen=500, 
    alpha=c(0.05,0.01), 
    keep.data = TRUE)

coolBlueHotRed <- function(n, alpha = 1) {
  rainbow(n, end=4/6, alpha=alpha)[n:1]
  }

pretty_palette <- c("#1f77b4","#ff7f0e","#2ca02c",
                    "#d62728","#9467bd","#8c564b","#e377c2")

plot(som_model, type = "property", property = getCodes(som_model)[,4], main=colnames(getCodes(som_model))[4], palette.name=coolBlueHotRed)
```

##Print out clusters by SOM
```{r}
cluster_som <- as.data.frame(allres$unit.classif)

cluster_som[,2] <- colnames(preprocess_clustering)
colnames(cluster_som) <- c("cluster", "phecode")

a <- lapply(c(1:25), function(i) {
  subset(cluster_som, cluster == i)
  })

split_cluster_som = list()
for (i in 1:length(a)) {
  split_cluster_som[[i]] <- merge(a[[i]],
                              as.data.frame(phecode_ukb[,1:4]), 
                              by = "phecode")
}

split_cluster_somdf <- data.table::rbindlist(split_cluster_som)

Heatmap(preprocess_clustering, name = "preprocess_clustering", 
        column_split = cluster_som[,1],
        cluster_rows = FALSE, col = col_fun)
```

##SOM clustering by 60 cut (2 ways)
```{r}
allres_60cut <- som(t(preprocess_clustering_60cut), 
                    grid = somgrid(5, 5))
plot(allres_60cut)

plot(allres_60cut, type = "counts", 
     palette.name = colors, heatkey = TRUE)

som_model_60cut <- som(t(preprocess_clustering_60cut), 
    grid=som_grid, 
    rlen=500, 
    alpha=c(0.05,0.01), 
    keep.data = TRUE)

plot(som_model_60cut, type = "property", 
     property = getCodes(som_model_60cut)[,4],
     main=colnames(getCodes(som_model_60cut))[4],
     palette.name=coolBlueHotRed)
```

##Print out clusters of 60 cut by SOM
```{r}
cluster_som_60cut <- as.data.frame(allres_60cut$unit.classif)

cluster_som_60cut[,2] <- colnames(preprocess_clustering_60cut)
colnames(cluster_som_60cut) <- c("cluster", "phecode")

a1 <- lapply(c(1:25), function(i) {
  subset(cluster_som_60cut, cluster == i)
  })

split_cluster_som_60cut = list()
for (i in 1:length(a)) {
  split_cluster_som_60cut[[i]] <- merge(a1[[i]],
                              as.data.frame(phecode_ukb[,1:4]), 
                              by = "phecode")
}

split_cluster_som_60cutdf <- data.table::rbindlist(split_cluster_som_60cut)

Heatmap(preprocess_clustering_60cut, 
        name = "preprocess_clustering_60cut", 
        column_split = cluster_som_60cut[,1],
        cluster_rows = FALSE, col = col_fun)
```

##Print out clusters using hclust and cutree and draw heatmap accordingly
```{r}
split = data.frame(cutree(hclust(dist(t(clust_result_2))), k = 8))

split[,2] <- rownames(split)
colnames(split) <- c("cluster", "phecode")

a <- lapply(c(1:8), function(i) {
  subset(split, cluster == i)
  })

split_cluster = list()
for (i in 1:length(a)) {
  split_cluster[[i]] <- merge(a[[i]],
                              as.data.frame(phecode_ukb[,1:4]), 
                              by = "phecode")
}

split_clusterdf <- data.table::rbindlist(split_cluster)

Heatmap(clust_result_2, name = "clust_result_2", 
        column_split = split[,1],
        cluster_rows = FALSE, col = col_fun)
```

##Drawing dataplotting_multiple by clusters
```{r}
source("Automated_functioncode.R")

phecode_cluster = list()
for (i in 1:length(split_cluster)) {
  phecode_cluster[[i]] <-
    multi_normalized_phecode(split_cluster[[i]]$phecode, i)
  print(i)
}

for (i in 1:length(phecode_cluster)) {
  dataplotting_multiple(phecode_cluster[[i]])
  print(i)
}
```

##Drawing phecode by categories
```{r}
unique_category <- unique(phecode_ukb$category)

category_phecode = list()
for (i in 1:length(unique_category)) {
  
  category_phecode[[i]] <- as.data.frame(subset(phecode_ukb, category ==
                                                  unique_category[i])[,1])
}

source("Automated_functioncode.R")

category_result = list()
for (i in 1:length(unique_category)) {
  category_result[[i]] <- multi_normalized_phecode(category_phecode[[i]]$phecode,
                                                   unique_category[i])
}

category_result[[3]][[2]] <- "endocrine_metabolic"
for (i in 13:length(category_result)) {
  dataplotting_multiple(category_result[[i]])
  print(i)
}
```
