---
title: "Automated_selfreportnoncancer"
author: "Jinhee Chang"
date: "10/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Self Reported Non Cancer automated

```{r}
# Load the data that I am going to use
selfreport_NC <- readRDS("SelfreportedNoncancer.rds")

# Check how this data looks like
head(selfreport_NC)

# Load Noncancercode and Cancercode data
Noncancercode <- readr::read_tsv("coding6.tsv")
Cancercode <- readr::read_tsv("coding3.tsv")
```

```{r}
## Don't forget to download this function
library(tidyr)
library(dplyr)
library(data.table)

sum(is.na(selfreport_NC$Noncancerage)) #None

# How many patient records are unique
length(unique(selfreport_NC$eid)) #379983 patients total
unique_selfreport_NC <- selfreport_NC[which(!duplicated(selfreport_NC$eid)), ]

# Check how this data looks like
head(unique_selfreport_NC)
```
## Check the age distribution of the total unique patients in selfreport_NC file

```{r}
hist(selfreport_NC$Noncancerage, breaks = 50)


# Check the minimum age and maximum age of the selfreport_NC
min(selfreport_NC$Noncancerage) # 0
max(selfreport_NC$Noncancerage) # 81
```

```{r}
# Don't forget to download this function
library(tidyr)
library(dplyr)
library(data.table)
library(ggplot2)

# code_num: Disease code from ICD10
test_selfreportNC <- function(code_num, freq = 100, age_min = 27, age_max = 65) {
  
  result <- selfreport_NC[grep(code_num, selfreport_NC$Noncancercode), ]
  
  if (nrow(result) < freq) {
    return(NULL)
  }
  
  age_freq <- as.data.frame(table(floor(result$Noncancerage)))
  colnames(age_freq) <- c("Age at Diagnosis", "Frequency")
  age_freq$`Incidence Rate` <- age_freq$Frequency/sum(age_freq$Frequency)
  age_freq$category <- rep(paste0(result$meaning[1], "(n=", nrow(result), ")"),
                           nrow(age_freq))

  age_freq$`Age at Diagnosis` <- as.numeric(as.character(age_freq$`Age at Diagnosis`))
  age_freq$category <- as.factor(age_freq$category)
  
  # Draw line plot
  plotdata<- ggplot(data = age_freq, aes(x = `Age at Diagnosis`,
                                  y = log10(`Incidence Rate`), color = category), 
                    las = 3)+
  #geom_line()+
  geom_smooth(alpha = .15, aes(fill = category))+
  scale_x_continuous(limits = c(age_min, age_max), n.breaks = 10)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(y = "log10(Incidence Rate)", x = "Age at Diagnosis")+
  theme_bw()+
  theme(legend.position = "right", 
        legend.direction = "vertical",
        legend.background = element_rect(color = "steelblue", linetype = "solid"),
        legend.text = element_text(size = 8))
  
    
  #print(plotdata)
  
  png(paste0(code_num, "_plot.png"), res = 80, width = 1000, height = 600)
  print(plotdata)
  dev.off() 
  
  return(age_freq)

}
```

## 1263 dementia/alzheimers/cognitive impairment (AD) - Total: 143

```{r}
# AD
test_selfreportNC("1263")
```

## 1076 heart failure/pulmonary odema (CHF) - Total: 358

```{r}
# CHF
test_selfreportNC("1076")
```

## 1112 chronic obstructive airways disease/copd - Total: 1982

```{r}
# COPD
test_selfreportNC("1112")
```
## 1075 heart attack/myocardial infarction - Total: 12142

```{r}
# Acute MI 
test_selfreportNC("1075")
```
## 1081 stroke - Total: 7021

```{r}
# Stroke
test_selfreportNC("1081")
```
## 1583 ischaemic stroke - Total: 32

```{r}
# Ischaemic Stroke
test_selfreportNC("1583")
```

## 1223 type 2 diabetes - Total: 4328

```{r}
# Diabetes Type II
test_selfreportNC("1223")
```

## 1220 diabetes - Total: 23000

```{r}
# Unspecified Diabetes 
test_selfreportNC("1220")
```
## negative control: 1567 infectious mononucleosis / glandular fever / epstein barr virus (ebv)

```{r}
test_selfreportNC("1567")
```

