---
title: "Script_phenotypes"
author: "Jinhee Chang"
date: "3/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(data.table)
library(foreach)
library(doParallel)
library(ukbtools)
library(dplyr)
library(purrr)
library(sampling)
library(moonBook)
library(R.utils)
library(pROC)
library(fmsb)
library(rsq)
library(reticulate)
```

### Covariate 1: Fid 21022 (Age at Recruitment)
```{r}
age = read.csv("21022.csv")
setnames(age, c("X21022.0.0"), c("age"))
```

### Covariate 2: Fid 31 (Sex)
```{r}
sex = read.csv("31.csv")
setnames(sex, c("X31.0.0"), c("sex"))

Pheno = merge(age[,c("eid","age")], sex[,c("eid","sex")],
              by = "eid", all.x = T, sort = F)
```

### Covariate 3: Fid 22000 (Genotype measurement batch (array_type))
```{r}
batch = read.csv("22000.csv")
setnames(batch, c("X22000.0.0"), c("batch"))

batch$array_type = NA
batch$array_type[!is.na(batch$batch) & batch$batch >= 1 & batch$batch <= 95] = 1
batch$array_type[!is.na(batch$batch) & batch$batch < 1] = 0
batch$array_type = as.factor(batch$array_type)
#length(batch$eid[batch$batch>=1 & batch$batch<=95])
#length(batch$eid[batch$array_type == 1])

Pheno = merge(Pheno, batch[,c("eid","array_type")],
            by = "eid", all.x = T, sort = F)
```

### Covariate 4: Fid 22009 (Genetic principal components (PCA))
```{r}
pca = read.csv("22009.csv")
col1 = paste0("X22009.0.", 1:20)
col2 = paste0("PC", 1:20)
setnames(pca, col1, col2)

Pheno = merge(Pheno, pca[,c("eid",col2)],
              by = "eid", all.x = T, sort = F)
```

### Covariate 5: Fid 21001 (Body mass index (BMI))

--> filled NA with the next available data for each row
--> first entry data for each patients were selected
```{r}
BMI = read.csv("21001.csv")

cat("BMI_firstcolumn_NA_count = ")
sum(is.na(BMI$X21001.0.0))
BMI$X21001.0.0[is.na(BMI$X21001.0.0)] <- BMI$X21001.1.0[is.na(BMI$X21001.0.0)]
cat("BMI_firstcolumn_NA_count = ")
sum(is.na(BMI$X21001.0.0))
BMI$X21001.0.0[is.na(BMI$X21001.0.0)] <- BMI$X21001.2.0[is.na(BMI$X21001.0.0)]
cat("BMI_firstcolumn_NA_count = ")
sum(is.na(BMI$X21001.0.0))
BMI$X21001.0.0[is.na(BMI$X21001.0.0)] <- BMI$X21001.3.0[is.na(BMI$X21001.0.0)]
cat("BMI_firstcolumn_NA_count = ")
sum(is.na(BMI$X21001.0.0))
cat("BMI_allcolumn_NA_count = ")
sum(rowSums(is.na(BMI)) == 4)

setnames(BMI, c("X21001.0.0"), c("BMI"))
Pheno = merge(Pheno, BMI[,c("eid","BMI")],
              by = "eid", all.x = T, sort = F)
```
### Exclude 1: Fid 22001 (Genetic sex) - sex mismatch 372 cases

exclude_eid_count = 372
```{r}
sex_mismatch=read.csv("22001.csv")
setnames(sex_mismatch, c("X22001.0.0"), c("genetic_sex"))

Pheno_temp = merge(sex[,c("eid","sex")], sex_mismatch[,c("eid","genetic_sex")],
                   by = "eid", all.x = T, sort = F)

exclude = Pheno_temp$eid[!is.na(Pheno_temp$genetic_sex) & !is.na(Pheno_temp$sex) & Pheno_temp$sex != Pheno_temp$genetic_sex]
exclude = exclude[!duplicated(exclude)]
cat("exclude_eid_count = ")
length(exclude) #372 mismatch
```
### Exclude 2: Fid 22019 (Sex chromosome aneuploidy) - 651 cases

sex_aneuploidy_not_NA_count = 651
Total_exclude_eid_count = 1023
Total_exclude_eid_count_nonduplicated = 842
```{r}
sex_aneuploidy = read.csv("22019.csv")
setnames(sex_aneuploidy, c("X22019.0.0"), c("aneuploidy_sex"))

cat("sex_aneuploidy_not_NA_count = ")
length(sex_aneuploidy$aneuploidy_sex[!is.na(sex_aneuploidy$aneuploidy_sex)]) #651 cases

exclude = c(exclude, sex_aneuploidy$eid[!is.na(sex_aneuploidy$aneuploidy_sex)])
cat("Total_exclude_eid_count = ")
length(exclude) #1023 cases

exclude = exclude[!duplicated(exclude)]
cat("Total_exclude_eid_count_nonduplicated = ")
length(exclude) #842 cases
```

### Exclude 3: Fid 22027 (Outliers for heterozygosity or missing rate) - 968 cases

het_missing_not_NA_count = 968
Total_exclude_eid_count = 1810
Total_exclude_eid_count_nonduplicated = 1805
```{r}
het_missing = read.csv("22027.csv")
setnames(het_missing, c("X22027.0.0"), c("het_missing"))

cat("het_missing_not_NA_count = ")
length(het_missing$het_missing[!is.na(het_missing$het_missing)]) #968 cases
exclude = c(exclude, het_missing$eid[!is.na(het_missing$het_missing)])
cat("Total_exclude_eid_count = ")
length(exclude) #1810 cases

exclude = exclude[!duplicated(exclude)]
cat("Total_exclude_eid_count_nonduplicated = ")
length(exclude) #1805 cases
```

### Exclude 4: Fid 22006 (Genetic ethnic grouping) - 92887 cases

non_white_british_count = 92887
Total_exclude_eid_count = 94692
Total_exclude_eid_count_nonduplicated = 94328
```{r}
white_british = read.csv("22006.csv")
setnames(white_british, c("X22006.0.0"), c("ethnicity_genetic"))

cat("non_white_british_count = ")
length(white_british$ethnicity_genetic[is.na(white_british$ethnicity_genetic)]) #92887 cases

non_white_british = white_british$eid[is.na(white_british$ethnicity_genetic)]
exclude = c(exclude, non_white_british)
cat("Total_exclude_eid_count = ")
length(exclude) #94692 cases

exclude = exclude[!duplicated(exclude)]
cat("Total_exclude_eid_count_nonduplicated = ")
length(exclude) #94328 cases
```

### Exclude 4: Fid 22021 (Genetic kinship to other participants) - 977 cases
```{r}
precal_kinship = read.csv("22021.csv")
setnames(precal_kinship, c("X22021.0.0"), c("kinship"))

# #-1: Participant excluded from kinship inference process
# precal_siblings = precal_kinship$eid[!is.na(precal_kinship$kinship) & precal_kinship$kinship == -1]
# length(precal_siblings[!duplicated(precal_siblings)]) #977
# 
# #0: No kinship found
# precal_siblings = precal_kinship$eid[!is.na(precal_kinship$kinship) & precal_kinship$kinship == 0]
# length(precal_siblings[!duplicated(precal_siblings)]) #339572
# 
# #1:	At least one relative identified  
# precal_siblings = precal_kinship$eid[!is.na(precal_kinship$kinship) & precal_kinship$kinship == 1]
# length(precal_siblings[!duplicated(precal_siblings)]) #147507
# 
# #10: Ten or more third-degree relatives identified 
# precal_siblings = precal_kinship$eid[!is.na(precal_kinship$kinship) & precal_kinship$kinship == 10]
# length(precal_siblings[!duplicated(precal_siblings)]) #188

exclude = c(exclude, precal_kinship$eid[!is.na(precal_kinship$kinship) & precal_kinship$kinship == -1])
cat("Total_exclude_eid_count = ")
length(exclude) #95305 cases
cat("Total_exclude_eid_count_nonduplicated = ")
length(exclude[!duplicated(exclude)]) #94330 cases

precal_siblings = precal_kinship$eid[!is.na(precal_kinship$kinship) & (precal_kinship$kinship %in% c(1,10))] #147695 cases
precal_siblings_multi=precal_kinship$eid[!is.na(precal_kinship$kinship) & (precal_kinship$kinship %in% c(10))] #188 cases
cat("precal_siblings_count = ")
length(precal_siblings[!duplicated(precal_siblings)])
cat("precal_siblings_multi_count = ")
length(precal_siblings_multi[!duplicated(precal_siblings_multi)])

for (chr in c(1:22, "MT", "X", "Y", "XY")) 
  {
  kinship_chr = fread(paste0("ukb32575_rel_chr", chr, "_s488264.dat")) #non overlapped data
  if (chr == 1) 
    {
    kinship = kinship_chr
  } else 
    {
    kinship = rbind(kinship, kinship_chr)
  }
}

kinship = kinship[kinship$ID1 %in% precal_siblings & kinship$ID2 %in% precal_siblings,]
kinship_nodup <- kinship %>% group_by(ID1,ID2) %>%
  summarize(Count = n())
kinship_nodup$Count = NULL

ids <- c(kinship_nodup$ID1, kinship_nodup$ID2) 

ids_dup = ids[duplicated(ids)] 
cat("Total_duplicated_ids_in_kinship = ")
length(ids_dup) #66562

ids_sin = ids[!duplicated(ids)] 
cat("Total_nonduplicated_ids_in_kinship = ")
length(ids_sin) #147652
not_founded_matched_ones <- setdiff(precal_siblings, ids) #43 cases -> exclude all (we cannot find matched persons in all ukb32575_rel_chr files)

exclude = c(exclude, not_founded_matched_ones)
exclude = exclude[!duplicated(exclude)]
cat("Total_exclude_eid_count_nonduplicated = ")
length(exclude) #94361 cases

kinship_nodup = kinship_nodup[!(kinship_nodup$ID1 %in% exclude) & !(kinship_nodup$ID2 %in% exclude),] #91215 cases
write.table(kinship_nodup, "kinship_matrix_with_relatedness.txt", col.names = T, row.names = F, quote = F, sep = "\t")

all_kinship = c(kinship_nodup$ID1, kinship_nodup$ID2)
all_kinship = all_kinship[!duplicated(all_kinship)]
cat("Total_all_kinship_count_nonduplicated = ")
length(all_kinship) #127542 cases
```
### Exclude 4 - Continued: Compare with pruning result

Total_exclude_eid_count_nonduplicated = 159075
```{r}
kinship_end = read.csv("kinship_matrix_with_relatedness_pruning.csv") #data from prune_subjects.py

kinship_exclude = setdiff(all_kinship, kinship_end$remaining_subjects) #64714 cases
exclude <- c(exclude, kinship_exclude)
exclude = exclude[!duplicated(exclude)]
cat("Total_exclude_eid_count_nonduplicated = ")
length(exclude) #159075 cases

exclude_df = data.frame(exclude_eid = exclude) 
write.table(exclude_df, "exclude_eid.txt", col.names = F, row.names = F, quote = F, sep = "\t")
```

### Write Pheno_final (Covariates with and without exclude eids)

--> Total_eid_count_without_exclude = 343411
--> Total_eid_count_with_exclude = 502486
```{r}
Pheno_final = Pheno[!(Pheno$eid %in% exclude_df$exclude_eid),]
cat("Total_eid_count_without_exclude = ")
length(Pheno_final$eid) #159075 cases

cat("Total_eid_count_with_exclude = ")
length(Pheno$eid) #159075 cases

write.table(Pheno, "Covariates_all.txt", col.names = T, row.names = F, quote = F, sep = "\t")
write.table(Pheno_final, "Covariates_noexclude.txt", col.names = T, row.names = F, quote = F, sep = "\t")
```
### Make table which eids in ukb_prune_data is included in covariates eids

TRUE: in covariates 
FALSE: excluded fromm covariates
```{r}
eid_in <- as.data.frame(table(ukb_prune_con$eid %in% Pheno_final$eid))
setnames(eid_in, c("Freq"), c("ukb_prune_con"))

temp <- as.data.frame(table(ukb_prune_lon$eid %in% Pheno_final$eid))
setnames(temp, c("Freq"), c("ukb_prune_lon"))
eid_in = merge(eid_in, temp[,c("Var1", "ukb_prune_lon")],
            by = "Var1", all.x = T, sort = F)

temp <- as.data.frame(table(ukb_prune_rep$eid %in% Pheno_final$eid))
setnames(temp, c("Freq"), c("ukb_prune_rep"))
eid_in = merge(eid_in, temp[,c("Var1", "ukb_prune_rep")],
            by = "Var1", all.x = T, sort = F)

temp <- as.data.frame(table(ukb_rep_lifespan$eid %in% Pheno_final$eid))
setnames(temp, c("Freq"), c("ukb_rep_lifespan"))
eid_in = merge(eid_in, temp[,c("Var1", "ukb_rep_lifespan")],
            by = "Var1", all.x = T, sort = F)

rownames(eid_in) <- eid_in[,1]
eid_in[,1] <- NULL

# write.table(eid_in, "eid_count_in_covariates.txt", col.names = T, row.names = T, quote = F, sep = "\t")
```

###Merging final ukb_rep_lifespan (without exclude data) with all other variables from ukb_prune_con, ukb_prune_lon, ukb_prune_rep and pheno variables (covariates)
```{r}
ukb_rep_lifespan_final = ukb_rep_lifespan[(ukb_rep_lifespan$eid %in% Pheno_final$eid),]

ukb_rep_lifespan_final = merge(ukb_rep_lifespan_final, ukb_prune_con,
                               by = "eid", all.x = T, sort = F)
ukb_rep_lifespan_final = merge(ukb_rep_lifespan_final, ukb_prune_lon,
                               by = "eid", all.x = T, sort = F)
ukb_rep_lifespan_final = merge(ukb_rep_lifespan_final, ukb_prune_rep,
                               by = "eid", all.x = T, sort = F)

ukb_rep_lifespan_final = merge(ukb_rep_lifespan_final, Pheno_final,
                               by = "eid", all.x = T, sort = F)

write.table(ukb_rep_lifespan_final, "ukb_rep_lifespan_final.txt", col.names = T, row.names = T, quote = F, sep = "\t")
```

###Make count_NA_by_var.txt that counts non-missing variables by each column in ukb_rep_lifespan_final
```{r}
count_NA_by_var <- apply(ukb_rep_lifespan_final, 2, 
                         function (x) sum(!is.na(x)))

write.table(count_NA_by_var, "count_NA_by_var.txt", col.names = T, row.names = T, quote = F, sep = "\t")
```

